---
title: "Modeling sleeping site choice with the basin of attractor model"
output:
html_document: default
html_notebook: default
pdf_document: default
---


##Introduction:

Here we use movement and sleeping site data to identify when and where we can preditct a groups final sleeping site. i.e., is there a basin of attraction around each sleeping site?

```{r, message=FALSE, warning=FALSE}
library(rstan)
library(ggplot2)
library(sp)
library(brms)
#library(raster)
#library(rgdal)
library(mgcv)
library(itsadug)
library(scales)
library(data.table)
library(dplyr)
#library("plot3Drgl")
#library(plot3D)
#library(rgl)
library(car)
library(scatterplot3d)
library(ggmap)
library(lubridate)
library(adehabitatLT)
```



```{r}
#write.csv(df.sub.sub,"obs_tracks_1h_sleepsite.csv", row.names = F)

#Read the data
tracks <- read.csv("data/obs_tracks_1h_sleepsite.csv")

#visualize the data
ggplot(tracks, aes(x=x,y=y,col=factor(sleepSite) )) + geom_point() 
```

```{r}

#model using multinomial
fit.sleepsite <- gam(list(sleepSite.target ~ te(x,y,t,bs=c("tp","tp"),k=c(25,5),d=c(2,1)),
                          ~ te(x,y,t,bs=c("tp","tp"),k=c(25,5),d=c(2,1)),
                          ~ te(x,y,t,bs=c("tp","tp"),k=c(25,5),d=c(2,1)),
                          ~ te(x,y,t,bs=c("tp","tp"),k=c(25,5),d=c(2,1))),
                     data=tracks, family = multinom(K=4))

summary(fit.sleepsite)
#plot(fit.sleepsite,pages=1)
gam.check(fit.sleepsite)
```


#Visualize the predictions

Predict for hourly blocks, and visualize in 3d
```{r}

#set spatio-temporal sampling
time.seq <- seq(6,17,by=0.2)
grd.all <- data.frame(x=-1,y=-1,t=-1)
for(i in time.seq){
  
  #Time
  time.pred <- i
  
  #create grid surrounding the observed points
  x.range <- as.numeric(c(min(tracks$x), max(tracks$x)))  
  y.range <- as.numeric(c(min(tracks$y), max(tracks$y) ))  
  
  #x.range <- as.numeric(c(78000, 80000))  
  #y.range <- as.numeric(c(18000, 22000 ))  
  grd <- expand.grid(x = seq(from = x.range[1], to = x.range[2], by = 100), y = seq(from = y.range[1], 
                                                                                   to = y.range[2], by = 100))  
  #What time slices to look at?
  grd$t <- time.pred
  
  #keep each time period
  grd.all <- rbind(grd.all, grd)
  
}
grd.all <- grd.all[-1,]


#prediction for each slice 
pp<-predict(fit.sleepsite, newdata = grd.all ,type="response", se.fit=T)

#get the predictions 
pc <- apply(pp$fit,1,function(x) which(max(x)==x)[1])

#remove the predictions with high uncertainty
pc.final <- vector()
pc.final.probs <- vector()
for(i in 1:length(pc)){
  pc.prob <- pp$fit[i,pc[i]]
  pc.se <- pp$se.fit[i,pc[i]]
  pc.final[length(pc.final)+1] <- ifelse(pc.prob - 1.96*pc.se > 0 & pc.prob>0.80 ,pc[i],NA)
  pc.final.probs[length(pc.final.probs)+1] <- pc.prob
}

#add predictions
grd.all$pc <- pc.final
grd.all$pc.prob <- pc.final.probs

#subset
grd.all.sub <- grd.all %>% filter(pc != 1)


#try exporting it as dynamic
library(scatterplot3d)
#library(knitr)
#library(rgl)
#knit_hooks$set(webgl = hook_webgl)
#scatter3d(x=grd.all.sub$x, y=grd.all.sub$y, z=grd.all.sub$t, groups = as.factor(grd.all.sub$pc), zlab = "time", xlab="x coord", ylab="y cood", surface = F)


###plot 3d
#graphics.off()
png(filename = "SleepSite_Basin_HR.png", width = 8, height = 6, units = "in", res = 300)
par(mai = c(0.5, 0.5, 0.5, 0.5))
scatterplot3d(x=grd.all.sub$x, y=grd.all.sub$y, z=grd.all.sub$t, color = c("yellow","green","blue","red", "orange")[grd.all.sub$pc], zlab = "time", xlab="x coord", ylab="y cood", surface = F, angle=20)
dev.off()


###export to shapefile
#library(rgdal)
#grd.all.sub$mgrs <- paste0("47QQU",round(grd.all.sub$x),round(grd.all.sub$y))
#lat.long<-mgrs_to_latlng(grd.all.sub$mgrs)
#to.arc <- cbind(grd.all.sub,lat.long)
#names(to.arc) <- c("MGRS_X","MGRS_Y","time","sleepSite","probS","mgrs1","mgrs2","lat", "long")
#write.csv(to.arc, file="sleepPred.csv")
#coordinates(to.arc)=c("long","lat")

#proj4string(to.arc)<- CRS("+proj=longlat +ellps=evrst30 ")
#proj4string(to.arc)<- CRS("+proj=longlat")
#raster::shapefile(LLcoor, "sleepSiteBasins3.shp", overwrite=T)

#to.arc@coords <- as.matrix(to.arc@data[c("Lon","Lat")])
#to.arc@bbox   <- rbind(range(to.arc@coords[,1]),range(to.arc@coords[,2]))
#library(ggmap)
#ggplotly(to.arc, aes(x=long, y=lat)) + geom_point()
#bas_map <- get_map(location=c(lon = mean(to.arc$long), lat=mean(to.arc$lat)), maptype = "satellite", zoom = 14)
#plottest<-get_map(location=c(lon = mean(to.arc$long), lat=mean(to.arc$lat)), maptype = "satellite", zoom = 14) %>% ggmap() +
#  geom_point(data = to.arc, aes(x = long, y = lat, col=factor(sleepSite) ), size = 3)

#ggplotly(plottest)

#ggplot(data = to.arc, aes(x = long, y = lat), color = 'red', size = 3) + geom_point() 

```



Visualize movement predictability 2d
```{r}

plot.12<-get_map(location=c(lon = mean(to.arc$long), lat=mean(to.arc$lat)), maptype = "satellite", zoom = 14) %>% ggmap() +
  geom_point(data = as.data.frame(to.arc[to.arc$time==12,]), aes(x = long, y = lat, col=factor(sleepSite) ), size = 3)+ theme(legend.position = "none")
plot.12

plot.12<-get_map(location=c(lon = mean(to.arc$long), lat=mean(to.arc$lat)), maptype = "satellite", zoom = 14) %>% ggmap() +
  geom_point(data = as.data.frame(to.arc[to.arc$time==12,]), aes(x = long, y = lat, col=factor(sleepSite) ), size = 3)+ theme(legend.position = "none")
plot.12


plot.13<-get_map(location=c(lon = mean(to.arc$long), lat=mean(to.arc$lat)), maptype = "satellite", zoom = 14) %>% ggmap() +
  geom_point(data = filter(to.arc, time==13), aes(x = long, y = lat, col=factor(sleepSite) ), size = 3)+ theme(legend.position = "none")

plot.14<-get_map(location=c(lon = mean(to.arc$long), lat=mean(to.arc$lat)), maptype = "satellite", zoom = 14) %>% ggmap() +
  geom_point(data = filter(to.arc, time==14), aes(x = long, y = lat, col=factor(sleepSite) ), size = 3)+ theme(legend.position = "none")

plot.15<-get_map(location=c(lon = mean(to.arc$long), lat=mean(to.arc$lat)), maptype = "satellite", zoom = 14) %>% ggmap() +
  geom_point(data = filter(to.arc, time==15), aes(x = long, y = lat, col=factor(sleepSite) ), size = 3)+ theme(legend.position = "none")

plot.16<-get_map(location=c(lon = mean(to.arc$long), lat=mean(to.arc$lat)), maptype = "satellite", zoom = 14) %>% ggmap() +
  geom_point(data = filter(to.arc, time==16), aes(x = long, y = lat, col=factor(sleepSite) ), size = 3)+ theme(legend.position = "none")

#cowplot::plot_grid(plot.13, plot.14, plot.15, plot.16) 
cowplot::plot_grid(plot.10, plot.12, plot.14, plot.16) 
```


plot of tracks used
```{r}
#sleep sites
df.sleepTrees <- as.data.frame(matrix(data=c(78756,78783,79628,80419,20445,20806,20838,21595), nrow=4, ncol=2))
names(df.sleepTrees)<-c("x","y")

#obs
tracks.com <- tracks[complete.cases(tracks),]
tracks.com$sleepSite.target2 <- tracks.com$sleepSite.target
tracks.com$sleepSite.target2 <- ifelse(tracks.com$sleepSite.target2==2,3,ifelse(tracks.com$sleepSite.target2==3,2,tracks.com$sleepSite.target2))
ggplot(tracks.com, aes(x=x,y=y,col=factor(sleepSite.target2) )) + geom_point() + geom_path(aes(group=date)) + scale_color_manual(values=c("grey40","green","red","blue", "orange") ) +theme_bw()+ geom_point(data=df.sleepTrees, aes(x=x,y=y), color="black",size=4) + theme(legend.position = "none")

#sim
ggplot(trak_with_sleep, aes(x=x,y=y, color=factor(sleepSite.target) )) + geom_point() + geom_path(aes(group=id))+ scale_color_manual(values=c("grey40","green","red","blue", "orange") ) + theme_bw()+
  geom_point(data=df.sleepTrees, color="black",size=4)+ theme(legend.position = "none")
```


plot of area using mesh
```{r plot mesh}

#set spatio-temporal sampling
time.seq <- seq(6,17,by=0.2)
grd.all <- data.frame(x=-1,y=-1,t=-1)
for(i in time.seq){
  
  #Time
  time.pred <- i
  
  #create grid surrounding the observed points
  x.range <- as.numeric(c(min(tracks$x), max(tracks$x)))  
  y.range <- as.numeric(c(min(tracks$y), max(tracks$y) ))  
  
  #x.range <- as.numeric(c(78000, 80000))  
  #y.range <- as.numeric(c(18000, 22000 ))  
  grd <- expand.grid(x = seq(from = x.range[1], to = x.range[2], by = 100), y = seq(from = y.range[1], 
                                                                                   to = y.range[2], by = 100))  
  #What time slices to look at?
  grd$t <- time.pred
  
  #keep each time period
  grd.all <- rbind(grd.all, grd)
  
}
grd.all <- grd.all[-1,]


#prediction for each slice 
pp<-predict(fit.sleepsite, newdata = grd.all ,type="response", se.fit=T)

#get the predictions 
pc <- apply(pp$fit,1,function(x) which(max(x)==x)[1])

#remove the predictions with high uncertainty
pc.final <- vector()
pc.final.probs <- vector()
for(i in 1:length(pc)){
  pc.prob <- pp$fit[i,pc[i]]
  pc.se <- pp$se.fit[i,pc[i]]
  pc.final[length(pc.final)+1] <- ifelse(pc.prob - 1.96*pc.se > 0 & pc.prob>0.80 ,pc[i],NA)
  pc.final.probs[length(pc.final.probs)+1] <- pc.prob
}

#add predictions
grd.all$pc <- pc.final
grd.all$pc.prob <- pc.final.probs

#subset
grd.all.sub <- grd.all %>% filter(pc != 1)

grd.all.sub$x <- as.numeric(scale(grd.all.sub$x))
grd.all.sub$y <- as.numeric(scale(grd.all.sub$y))
grd.all.sub$t <- as.numeric(scale(grd.all.sub$t))


library(rgl)
library(misc3d)
bunny.dens <- kde3d(grd.all.sub[,"x"],grd.all.sub[,"y"],grd.all.sub[,"t"], n=150) # I chose lim values manually

plot(bunny.dens)
contour3d(bunny.dens$d, level = 0.08, 
    color = "green", color2 = "green", smooth=500)



```



#Simulate using empirically fitted corelated bias random movements

###Outline
#####1. fit a circular cirular model to the tracks (predicting both distance traveled and angle of travel, as well as their correlation).

#####2. simulate data using circular circular model, with a bias towards the nearest sleep site after 15:00. Stop sim at 18:00

#####3. model the basin of attractors that result.

#####4. take the difference in those 2d plots to highlight the differences in space and time.

#####5. Are there shared characteristics between the sites?

Calculate step lengths and turn angles
```{r}

library(adehabitatLT)
library(sp)

#get tracks data into spatial data
merge <- tracks[complete.cases(tracks),]
merge$date_time <- lubridate::dmy(merge$date) + lubridate::hours(floor(merge$t) ) + lubridate::minutes( round((merge$t-floor(merge$t))*60) )
coordinates(merge) <- ~x + y

ltraj <- as.ltraj(coordinates(merge),date=as.POSIXct(merge$date_time),id=1)

#take a look
plot(ltraj)
plotltr(ltraj, "dt/3600")

ltraj.sub<-cutltraj(ltraj, "dt > 3600+60*30")
plot(ltraj.sub)

burst.data<-ld(ltraj.sub)
burst.data
```

Get the data ready for stan
```{r}

#get as dataframe
df.angle<-burst.data

#get the previous steps measures
df.angle <- df.angle %>% group_by(id) %>% mutate(prev.angle = lag(abs.angle), prev.dist = lag(dist),prev.t.dist = lag(dt))

#keep only complete cases
df.angle<-df.angle[complete.cases(df.angle),]

#fix the time
#df.angle[df.angle$prev.t.dist>1.5*60*60,]<-NA
#df.angle<-df.angle[complete.cases(df.angle),]

#take a look
summary(df.angle$dist)
summary(df.angle$prev.dist)  


```


```{r}
data.rm<-list( Ni =  nrow(df.angle),
               Nj = length(unique(df.angle$id)),
               id = as.numeric(as.character(df.angle$id)),
               a = df.angle$abs.angle,
               pa = df.angle$prev.angle,
               d = as.numeric(scale(df.angle$dist, center = F)),
               pd = as.numeric(scale(df.angle$prev.dist, center = F)),
               dt = as.numeric( scale(df.angle$dt, center=F) )
)

```

The correlated random walk model (autocorrelated N and theta)
```{r}

model.rm = '

data {
    int Ni;          //Number of observations
    int Nj;          //Number of individuals
    vector [Ni] id;  //Individual id

    vector [Ni] a;   //Observed angle
    vector [Ni] pa;  //previous angle
    vector [Ni] d;   //distance traveled
    vector [Ni] pd;  //previous distance traveled
    vector [Ni] dt;  //time since last obs
}

transformed data {

vector[Ni] py;
vector[Ni] px;

for(i in 1:Ni){

 py[i] = sin(pa[i]);
 px[i] = cos(pa[i]);

}

}

parameters {
    unit_vector [2] mu_vec;
    real <lower = 0> kappa;
    real b_a;
    real mu_d;
    real b_d;
    real b_t;
    real<lower=0> sigma;
}

transformed parameters {

    vector[Ni] a_hat;
    vector[Ni] d_hat;

    d_hat = mu_d + b_d*pd + b_t*dt;

    for(i in 1:Ni){
          a_hat[i] = atan2(mu_vec[2] + b_a*py[i], mu_vec[1] + b_a*px[i]);
    }
}

model {
    kappa ~  normal (0, 1);
    sigma ~  normal (0, 1);
    b_a ~ normal(0,1);
    b_d ~ normal(0,1);
    mu_d ~ normal(0,1);

    a ~ von_mises (a_hat, kappa);
    d ~ lognormal(d_hat, sigma);
}

'
```


data.rm
```{r}
fit.rm <- stan(model_code = model.rm, data = data.rm, iter=500, chains=4, cores=4, init_r=0.1, control=list(adapt_delta=0.99, max_treedepth=12))
```


```{r}
print(fit.rm, pars = c("b_a","kappa","b_d","mu_d","b_t", "sigma"))
#pairs(fit.rm, pars = c("b_a","kappa","b_d","mu_d", "sigma"))
#traceplot(fit.rm, pars = c("b_a","kappa","b_d","mu_d", "sigma"))
```


#2) simulate some data to compare

Simulated data to make predictions
```{r}
library(brms)

post.rm<-rstan::extract(fit.rm)

sim_one_day <-function(x_start,y_start, post=post.rm){
  
  #back dist
  back.dist.prev <- sd(df.angle$prev.dist)
  back.dist.obs <- sd(df.angle$dist)
  
  #starting direction and distance
  a_start = rvon_mises(1,atan2(mean(post$mu_vec[2]),mean(post$mu_vec[1]) ),0.01 )
  d_start = min(rlnorm(1, mean(post$mu_d), mean(post$sigma) ) * back.dist.obs,max(df.angle$dist) )
  
  #sample a trajectory length and start time
  unique.track <- unique(tracks$date)
  start.df <-tracks %>% filter(date == unique.track[sample(1:length(unique.track),1)] )
  #start.time<-hour(round_date( start.df[1,]$date, "hour"))
  start.time<-round(start.df$t)[1]
  numb_setps = max(nrow(start.df),2)
  
  #start of sim data
  sim.data <- data.frame(x=x_start,y=y_start,time=start.time,id=1, a=a_start,d=d_start   )
    
  #simulate numb_setps
  for( i in 2:numb_setps){
    
    post_r <- sample(1:nrow(post$mu_vec),1)
    
    mean_a <- atan2(post$mu_vec[post_r,2] + post$b_a[post_r] * sin(sim.data$a[i-1]), post$mu_vec[post_r,1] + post$b_a[post_r] * cos(sim.data$a[i-1]) )  
    
    mean_d <- post$mu_d[post_r] + post$b_d[post_r]*(sim.data$d[i-1]/back.dist.prev) + post$b_t[post_r]*1
    
    obs_angle <- rvon_mises(1,mean_a, post$kappa[post_r]) 
    obs_dist  <- (rlnorm(1,mean_d, post$sigma[post_r]))*back.dist.obs
    
    #update x and y
    x = sim.data$x[i-1] + cos(obs_angle)*obs_dist
    y = sim.data$y[i-1] + sin(obs_angle)*obs_dist
    
    t = sim.data$time[i-1] + 1
    
    #Join to dataframe and continue
    df.temp <- data.frame(x=x,y=y,time=t, id=1,a=obs_angle,d=obs_dist)
    sim.data <- rbind(sim.data,df.temp)
    
    #ggplot(sim.data, aes(x=x,y=y)) + geom_line()
    
  }

  return (sim.data)
  
  
  
}
```


Just a test
```{r}
for(i in 1:1000){
  df.sim1 <- sim_one_day(x=sleepingSite[1], y=sleepingSite[2])
}
```


Run the same number of observed tracks with the simulation model
```{r}
#setup data frame to store data
trak_with_sleep <- data.frame(x=-1,y=-1, time=-1,id=1,a=-1,d=-1, sleepsite=-1)
track.id = 1

#set sleeping sites to choose from
actualSleepingSites <- matrix(data=c(78756,78783,79628,80419,20445,20806,20838,21595), nrow=4, ncol=2)

#randomly start at one sleeping site
sleepingSite <- actualSleepingSites[sample(1:4,1),]

#simulate movement
for(i in 1:1092){      #length(unique(tracks$date))...sum(numb.tracks.per.sleepsite[,2])
  
  #simulate one movement
  df.sim1 <- sim_one_day(x=sleepingSite[1], y=sleepingSite[2])
  x.last<-df.sim1[nrow(df.sim1),]$x
  y.last<-df.sim1[nrow(df.sim1),]$y
  
  #found sleep site nearby the end of the track
  retval = FALSE
  dist.min=500
  siteChoosen = 0
  
  #look for closest sleeping site from the end of the track
  for(j in 1:nrow(actualSleepingSites) ){
    
    #calculate distance to all sites  
    dist.temp <- (((actualSleepingSites[j,1]-x.last)^2) + ((actualSleepingSites[j,2]-y.last)^2) )^0.5
    
    #if the track is within 500m of the j sleeping site
    if(dist.temp < 500 & dist.min > dist.temp ){
      
      retval = TRUE
      siteChoosen = j
      dist.min=dist.temp
      
    } 
  }
  
  #no sleep site was found
  if(retval==FALSE){
    df.sim1$sleepsite = 0
    df.sim1$id=track.id
    track.id = track.id + 1
    trak_with_sleep <- rbind(trak_with_sleep, df.sim1)
    print(track.id)
    print(paste0("sleepsite = 0, at a dist = ",dist.temp))
    
    #random sleep site
    sleepingSite <- actualSleepingSites[sample(1:4,1),]
    
    #yes sleep site found
  } else {
    
    df.sim1$sleepsite = siteChoosen
    df.sim1$id=track.id
    track.id = track.id + 1
    trak_with_sleep <- rbind(trak_with_sleep, df.sim1)
    print(track.id)
    print(paste0("sleepsite = ",siteChoosen," at a dist = ",dist.temp))
    
    #change sleep sites
    sleepingSite <- actualSleepingSites[siteChoosen,]
  }
}


trak_with_sleep<- trak_with_sleep[-1,]
names(trak_with_sleep)[3] <- "t"
names(trak_with_sleep)[7] <- "sleepSite.target"

df.sleepTrees <- as.data.frame(matrix(data=c(78756,78783,79628,80419,20445,20806,20838,21595), nrow=4, ncol=2))
names(df.sleepTrees)<-c("x","y")
ggplot(trak_with_sleep, aes(x=x,y=y, color=id)) + geom_point() + geom_path() + theme_classic()+
  geom_point(data=df.sleepTrees, color="red",size=2)
```


add background prior 
```{r}
#add background priors
x.range.na <- as.numeric(c(min(trak_with_sleep$x)-50, max(trak_with_sleep$x)+50))  
y.range.na <- as.numeric(c(min(trak_with_sleep$y)-50, max(trak_with_sleep$y)+50))  
grd <- expand.grid(x = seq(from = x.range.na[1], to = x.range.na[2], by =100), y = seq(from = y.range.na[1], 
                                                                                               to = y.range.na[2], by = 100))  
#What time slices to look at?
grd$t <- runif(nrow(grd),min=4,max=19) #little larger than the actual time

grd.df <- data.frame(x=grd$x,y=grd$y,t=grd$t, id=0,a=-1,d=-1,sleepSite.target=0)
trak_with_sleep.grid <- rbind(trak_with_sleep,grd.df)

table(trak_with_sleep.grid$sleepSite.target)

#plot tracks coloured by where the troop ends up at night
ggplot(trak_with_sleep.grid, aes(x=x,y=y,col=factor(sleepSite.target) )) + geom_point() 

```

Comapre sample sizes
```{r}
#observed
tracks[-c( (nrow(tracks)-1960):nrow(tracks)),]%>% group_by(sleepSite.target) %>% tally()
sum((tracks[-c( (nrow(tracks)-1960):nrow(tracks)),] %>% group_by(sleepSite.target) %>% tally())$n)


#simulated
trak_with_sleep %>% group_by(sleepSite.target) %>% tally()
sum((trak_with_sleep %>% group_by(sleepSite.target) %>% tally())$n)
```



#MODEL BASINS
```{r}

trak_with_sleep.grid %>% group_by(sleepSite.target) %>% tally()

#model using multinomial
fit.sleepsite.sim <- gam(list(sleepSite.target ~ te(x,y,t,bs=c("tp","tp"),k=c(25,5),d=c(2,1)),
                          ~ te(x,y,t,bs=c("tp","tp"),k=c(25,5),d=c(2,1)),
                          ~ te(x,y,t,bs=c("tp","tp"),k=c(25,5),d=c(2,1)),
                          ~ te(x,y,t,bs=c("tp","tp"),k=c(25,5),d=c(2,1))),
                     data=trak_with_sleep.grid, family = multinom(K=4))

summary(fit.sleepsite.sim)
#plot(fit.sleepsite,pages=1)
gam.check(fit.sleepsite.sim)
```





#Visualize the predictions

Predict for hourly blocks, and visualize in 3d

```{r}

#prediction for each slice 
pp.sim<-predict(fit.sleepsite.sim, newdata = grd.all ,type="response", se.fit=T)

#get the predictions 
pc.sim <- apply(pp.sim$fit,1,function(x) which(max(x)==x)[1])

#remove the predictions with high uncertainty
pc.final.sim <- vector()
pc.final.probs.sim <- vector()
for(i in 1:length(pc.sim)){
  pc.prob <- pp.sim$fit[i,pc.sim[i]]
  pc.se <- pp.sim$se.fit[i,pc.sim[i]]
  pc.final.sim[length(pc.final.sim)+1] <- ifelse(pc.prob - 1.96*pc.se > 0 & pc.prob>0.80 ,pc.sim[i],NA)
  pc.final.probs.sim[length(pc.final.probs.sim)+1] <- pc.prob
}

#add predictions
grd.all$pc.sim <- pc.final.sim
grd.all$pc.prob.sim <- pc.final.probs.sim

#subset
grd.all.sub.sim <- grd.all %>% filter(pc.sim != 1)

####try exporting it as dynamic
#library(knitr)
#library(rgl)
#knit_hooks$set(webgl = hook_webgl)
#scatter3d(x=grd.all.sub.sim$x, y=grd.all.sub.sim$y, z=grd.all.sub.sim$t, groups = as.factor(grd.all.sub.sim$pc.sim), zlab = "time", xlab="x coord", ylab="y cood", surface = F)

###plot 3d
library(scatterplot3d)
#graphics.off()
png(filename = "SleepSite_Basin_sim.png", width = 8, height = 6, units = "in", res = 300)
par(mai = c(0.5, 0.5, 0.5, 0.5))
scatterplot3d(x=grd.all.sub.sim$x, y=grd.all.sub.sim$y, z=grd.all.sub.sim$t, color = c("yellow","green","red","blue", "orange")[grd.all.sub.sim$pc], zlab = "time", xlab="x coord", ylab="y cood", surface = F, angle=20, zlim=c(8,18))
dev.off()

#export to shapefile
#library(rgdal)
#grd.all.sub$mgrs <- paste0("47QQU",grd.all.sub$x,grd.all.sub$y)
#lat.long<-mgrs_to_latlng(grd.all.sub$mgrs)
#to.arc <- cbind(grd.all.sub,lat.long)
#names(to.arc) <- c("MGRS_X","MGRS_Y","time","sleepSite","probS","mgrs1","mgrs2","lat", "long")
#write.csv(to.arc, file="sleepPred.csv")
#coordinates(to.arc)=~lat+long

#proj4string(to.arc)<- CRS("+proj=longlat +ellps=evrst30 ")
#proj4string(to.arc)<- CRS("+proj=longlat")
#raster::shapefile(LLcoor, "sleepSiteBasins2.shp", overwrite=T)

#to.arc@coords <- as.matrix(to.arc@data[c("Lon","Lat")])
#to.arc@bbox   <- rbind(range(to.arc@coords[,1]),range(to.arc@coords[,2]))
#library(ggmap)
#ggplotly(to.arc, aes(x=long, y=lat)) + geom_point()
#bas_map <- get_map(location=c(lon = mean(to.arc$long), lat=mean(to.arc$lat)), maptype = "satellite", zoom = 14)
#plottest<-get_map(location=c(lon = mean(to.arc$long), lat=mean(to.arc$lat)), maptype = "satellite", zoom = 14) %>% ggmap() +
#  geom_point(data = to.arc, aes(x = long, y = lat, col=factor(sleepSite) ), size = 3)

#ggplotly(plottest)

#ggplot(data = to.arc, aes(x = long, y = lat), color = 'red', size = 3) + geom_point() 

```

Plot the difference between the two basins
```{r}

grd.all.diff <- grd.all

#calculate difference in probability
grd.all.diff$diff_prob <- grd.all.diff$pc.prob- grd.all.diff$pc.prob.sim

#calulate difference in classes
grd.all.diff$pc0 <- ifelse(is.na(grd.all.diff$pc),-99,grd.all.diff$pc)
grd.all.diff$pc.sim0 <- ifelse(is.na(grd.all.diff$pc.sim),-99,grd.all.diff$pc.sim)
grd.all.diff$diff_cat_na <- ifelse(grd.all.diff$pc0 == grd.all.diff$pc.sim0,grd.all.diff$pc.sim0,paste0(grd.all.diff$pc.sim0,"-",grd.all.diff$pc0))

table(grd.all.diff$diff_cat)
table(grd.all.diff$diff_cat_na)


#table(grd.all.diff.sub$diff_cat_na)
#levels(as.factor(grd.all.diff.sub$diff_cat_na))

#greater/less than random (some diff)
grd.all.diff.sub <- grd.all.diff %>% filter(diff_cat_na != "1", diff_cat_na!="-99", diff_cat_na!="1--99",diff_cat_na!="-99-1")

#recode
grd.all.diff.sub$gtlt.random <- grd.all.diff.sub$diff_cat_na
grd.all.diff.sub$gtlt.random[grd.all.diff.sub$gtlt.random=="-99-2"] <- "gt"
grd.all.diff.sub$gtlt.random[grd.all.diff.sub$gtlt.random=="-99-3"] <- "gt"
grd.all.diff.sub$gtlt.random[grd.all.diff.sub$gtlt.random=="-99-5"] <- "gt"
grd.all.diff.sub$gtlt.random[grd.all.diff.sub$gtlt.random=="1-2"] <- "gt"
grd.all.diff.sub$gtlt.random[grd.all.diff.sub$gtlt.random=="1-5"] <- "gt"
grd.all.diff.sub$gtlt.random[grd.all.diff.sub$gtlt.random=="2--99"] <- "lt"
grd.all.diff.sub$gtlt.random[grd.all.diff.sub$gtlt.random=="3--99"] <- "lt"
grd.all.diff.sub$gtlt.random[grd.all.diff.sub$gtlt.random=="4--99"] <- "lt"
grd.all.diff.sub$gtlt.random[grd.all.diff.sub$gtlt.random=="5--99"] <- "lt"
grd.all.diff.sub$gtlt.random[grd.all.diff.sub$gtlt.random=="5-1"] <- "lt"
grd.all.diff.sub$gtlt.random[grd.all.diff.sub$gtlt.random=="5"] <- "same"
grd.all.diff.sub$gtlt.random[grd.all.diff.sub$gtlt.random=="2"] <- "same"
grd.all.diff.sub$gtlt.random[grd.all.diff.sub$gtlt.random=="3-4"] <- "same"
grd.all.diff.sub$gtlt.random[grd.all.diff.sub$gtlt.random=="4-3"] <- "same"

table(grd.all.diff.sub$gtlt.random)

scatter3d(x=grd.all.diff.sub$x, y=grd.all.diff.sub$y, z=grd.all.diff.sub$t, groups = as.factor(grd.all.diff.sub$gtlt.random), zlab = "time", xlab="x coord", ylab="y cood", surface = F,surface.col=c("red","blue","grey50"))

# png(filename = "SleepSite_Basin_sim_diff.png", width = 8, height = 6, units = "in", res = 300)
# par(mai = c(0.5, 0.5, 0.5, 0.5))
# scatterplot3d(x=grd.all.diff.sub$x, y=grd.all.diff.sub$y, z=grd.all.diff.sub$t, color = c("red","blue","grey50")[as.numeric(as.factor(grd.all.diff.sub$gtlt.random))], zlab = "time", xlab="x coord", ylab="y cood", surface = F, angle=20, zlim=c(8,18))
# dev.off()
```



```{r}
grd.all.diff.sub$mgrs <- paste0("47QQU",round(grd.all.diff.sub$x),round(grd.all.diff.sub$y))
lat.long<-mgrs_to_latlng(grd.all.diff.sub$mgrs)
to.arc.diff <- cbind(grd.all.diff.sub,lat.long)
names(to.arc.diff) <- c("MGRS_X","MGRS_Y","time","sleepSite","probS","pc.sim","pc.prob.sim","diff.prob","pc0","pc.sim0","diff_cat","gtlt.random","mgrs","mgrs0","lat","long")
#write.csv(to.arc, file="sleepPred.csv")
coordinates(to.arc.diff)=c("long","lat")

proj4string(to.arc.diff)<- CRS("+proj=longlat +ellps=evrst30 ")

#convert sleepsite coords
df.sleepTrees <- as.data.frame(matrix(data=c(78756,78783,79628,80419,20445,20806,20838,21595), nrow=4, ncol=2))
df.sleepTrees$mgrs <- paste0("47QQU",round(df.sleepTrees[,1]),round(df.sleepTrees[,2]))
lat.long<-mgrs_to_latlng(df.sleepTrees$mgrs)
df.sleepTrees<-cbind(df.sleepTrees,lat.long)

```

Visualize in 2d
```{r}
library(ggmap)

plot.diff.11<-get_map(location=c(lon = mean(to.arc.diff$long)+0.002, lat=mean(to.arc.diff$lat))+0.001, maptype = "satellite", zoom = 15) %>% ggmap() + geom_point(data = as.data.frame(to.arc.diff[to.arc.diff$time==11,]), aes(x = long, y = lat, color=as.data.frame(to.arc.diff[to.arc.diff$time==11,])$gtlt.random), size = 2,shape=1)+
  geom_point(data=lat.long, aes(x = lng, y = lat),color="white", size=4) + scale_color_manual(values=c("orange", "#56B4E9","grey90" ))+ theme(legend.position = "none")

#plot.diff.11

plot.diff.13<-get_map(location=c(lon = mean(to.arc.diff$long)+0.002, lat=mean(to.arc.diff$lat))+0.001, maptype = "satellite", zoom = 15) %>% ggmap() + geom_point(data = as.data.frame(to.arc.diff[to.arc.diff$time==13,]), aes(x = long, y = lat, color=as.data.frame(to.arc.diff[to.arc.diff$time==13,])$gtlt.random), size = 2,shape=1)+
  geom_point(data=lat.long, aes(x = lng, y = lat),color="white", size=4)+ scale_color_manual(values=c("orange", "#56B4E9","grey90" ))+ theme(legend.position = "none")
#plot.diff.13

plot.diff.15<-get_map(location=c(lon = mean(to.arc.diff$long)+0.002, lat=mean(to.arc.diff$lat))+0.001, maptype = "satellite", zoom = 15) %>% ggmap() + geom_point(data = as.data.frame(to.arc.diff[to.arc.diff$time==15,]), aes(x = long, y = lat, color=as.data.frame(to.arc.diff[to.arc.diff$time==15,])$gtlt.random), size = 2,shape=1)+
  geom_point(data=lat.long, aes(x = lng, y = lat),color="white", size=4)+ scale_color_manual(values=c("orange", "#56B4E9","grey90" ))+ theme(legend.position = "none") 


plot.diff.17<-get_map(location=c(lon = mean(to.arc.diff$long)+0.002, lat=mean(to.arc.diff$lat))+0.001, maptype = "satellite", zoom = 15) %>% ggmap() + geom_point(data = as.data.frame(to.arc.diff[to.arc.diff$time==17,]), aes(x = long, y = lat, color=as.data.frame(to.arc.diff[to.arc.diff$time==17,])$gtlt.random), size = 2,shape=1)+
  geom_point(data=lat.long, aes(x = lng, y = lat),color="white", size=4)+ scale_color_manual(values=c("orange", "#56B4E9","grey90" ))+ theme(legend.position = "none") 

#plot.diff.17

cowplot::plot_grid(plot.diff.11,plot.diff.13,plot.diff.15,plot.diff.17, ncol=2)

```
